Characterizing Stellar Streams by ML

### 2023-08-29 src/tex/ms.tex
[showyourwork] overleaf sync

        This defines an astrometric phase-space hyper-tube
        \cite{2018MNRAS.477.4063M}, with a distribution radial to the track.

@@

        \begin{enumerate}
            \item Streamfinder \citep{2018MNRAS.477.4063M}. This method requires the potential
            \item VIA MACHINAE \citep{2022MNRAS.509.5992S}
            \item Kiyan and Adrian
            \item Matched Filter
            \item Gaia stuff: Adrian and Ana and Sara et al
            \item Uniform Modeling of 12 Stellar stream \citep{PatrickEtAl2022}. Only on phi1, phi2, and CMD. Their CMD method is more sophisticated. Worth discussing.
        \end{enumerate}

@@

        Streams have a diversity of measurement coverage: several have full 6-D
        astrometrics as well as photometrics and chemical abundances available
        \citep[e.g.,][]{Koposov2019, 2020A&A...635L...3A, 2022ApJ...928...30L},
        while most have partial coverage, lacking some feature dimensions.
        % TODO: terminology ND-phase-space vs feature columns
        Most streams in \Gaia {\color{red} [cite]}, for instance, have proper
        motions but not radial velocities nor reliable parallax measurements.
        Followup surveys, like \dataarchive{S5} {\color{red} [cite]}, add many
        feature dimensions, however only for a select set of streams. Given the
        diversity of coverage it is important that the likelihood model --
        \autoref{eq:membership_prob} -- account for missing data.

@@

            % TODO: rephrase
            By modeling jointly the astrometrics and photometrics we may tie
            together components of the models.  In particular, the distance
            modulus in the photometric model and the parallax in the astrometric
            model are both measures of the distance track of the stream.  We
            introduce a prior to connect these components of the model:
            \begin{align}
                \prior(\{\mu_I, \mu_\parallax \} | \phi_1) &= \delta( \rm{dist}(\mu_I) - \rm{dist}(\parallax) )(\phi_1), \label{eq:distance_track_prior}
                % \prior(\{\sigma_I, \sigma_\parallax \}(\phi_1)) &= \delta( \rm{dist}(\mu_I) - \rm{dist}(\parallax) )(\phi_1) \label{eq:distance_width_prior}
            \end{align}
            % TODO: should we show that these are different distance functions? or is it obvious
            where $\delta$ is the Dirac delta function
            \citep{1947pqm..book.....D}, and enforces that the distance modulus
            track and parallax track must be equal (when converted to the actual
            distance) as a function of $\phi_1$.  The widths are similarly
            constrained. The distance and first-order error conversion are given
            by
            \begin{align}
                \mu_I(\parallax) &= 10 - 5\log_{10}(\parallax) \\
                \sigma_I(\sigma_\parallax)  &\approxeq \frac{5}{\ln{10}} \mu_I \lvert\sigma_\parallax\rvert.
            \end{align}

@@

            For each parameter in $\mbs{\theta}$ of the model we implement
            priors on its range.  Uninformative, e.g. uniform $\mcal{U}(a, b)$,
            priors are often assumed where little is known about a parameter
            (except that $(a,b]$ contain the correct $\theta$ and its
            distribution). Where more is known about $\theta$, for instance that
            it is a categorical, more informative priors like the Dirichlet
            distribution \citep{Bishop_Christopher_M_2016-08-23} may be used.
            Our code permits any user-defined prior, and we include $\mcal{U}(a,
            b)$.  The default choice of prior and bounds $\mcal{U}$ set to the
            limits of the dataset in each feature dimension.
            % ⬆ where should this sentence be?

@@

        \subsubsection[STREAMFINDER]{\cite{2018MNRAS.477.4063M}}

>>>>

        This defines an astrometric phase-space hyper-tube
        \cite{STREAMFINDER}, with a distribution radial to the track.

@@

        \begin{enumerate}
            \item Streamfinder \citep{STREAMFINDER}. This method requires the potential
            \item VIA MACHINAE \citep{2022MNRAS.509.5992S}
            \item Kiyan and Adrian
            \item Matched Filter
            \item Gaia stuff: Adrian and Ana and Sara et al
            \item Uniform Modeling of 12 Stellar stream \citep{PatrickEtAl2022}. Only on phi1, phi2, and CMD. Their CMD method is more sophisticated. Worth discussing.
        \end{enumerate}

@@

        Streams have a diversity of measurement coverage: several have full 6-D
        astrometrics as well as photometrics and chemical abundances available
        \citep[e.g.,][]{Koposov2019, Antoja2020, 2022ApJ...928...30L},
        while most have partial coverage, lacking some feature dimensions.
        % TODO: terminology ND-phase-space vs feature columns
        Most streams in \Gaia {\color{red} [cite]}, for instance, have proper
        motions but not radial velocities nor reliable parallax measurements.
        Followup surveys, like \dataarchive{S5} {\color{red} [cite]}, add many
        feature dimensions, however only for a select set of streams. Given the
        diversity of coverage it is important that the likelihood model --
        \autoref{eq:membership_prob} -- account for missing data.

@@

            % TODO: rephrase
            By modeling jointly the astrometrics and photometrics we may tie
            together components of the models.  In particular, the distance
            modulus in the photometric model and the parallax in the astrometric
            model are both measures of the distance track of the stream.  We
            introduce a prior to connect these components of the model:
            \begin{align}
                \prior(\{\mu_I, \mu_\parallax \} | \phi_1) &= \delta( \rm{dist}(\mu_I) - \rm{dist}(\parallax) )(\phi_1), \label{eq:distance_track_prior}
                % \prior(\{\sigma_I, \sigma_\parallax \}(\phi_1)) &= \delta( \rm{dist}(\mu_I) - \rm{dist}(\parallax) )(\phi_1) \label{eq:distance_width_prior}
            \end{align}
            % TODO: should we show that these are different distance functions? or is it obvious
            where $\delta$ is the Dirac delta function
            \citep{Dirac1947}, and enforces that the distance modulus
            track and parallax track must be equal (when converted to the actual
            distance) as a function of $\phi_1$.  The widths are similarly
            constrained. The distance and first-order error conversion are given
            by
            \begin{align}
                \mu_I(\parallax) &= 10 - 5\log_{10}(\parallax) \\
                \sigma_I(\sigma_\parallax)  &\approxeq \frac{5}{\ln{10}} \mu_I \lvert\sigma_\parallax\rvert.
            \end{align}

@@

            For each parameter in $\mbs{\theta}$ of the model we implement
            priors on its range.  Uninformative, e.g. uniform $\mcal{U}(a, b)$,
            priors are often assumed where little is known about a parameter
            (except that $(a,b]$ contain the correct $\theta$ and its
            distribution). Where more is known about $\theta$, for instance that
            it is a categorical, more informative priors like the Dirichlet
            distribution \citep{Christopher2016} may be used.
            Our code permits any user-defined prior, and we include $\mcal{U}(a,
            b)$.  The default choice of prior and bounds $\mcal{U}$ set to the
            limits of the dataset in each feature dimension.
            % ⬆ where should this sentence be?

@@

        \subsubsection[STREAMFINDER]{\cite{STREAMFINDER}}
